---
title: "Getting Started with ultragsea"
author: "BigOmics Analytics"
package: ultragsea
output:
    BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Getting Started with ultragsea}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

# Introduction

ultragsea is a novel, ultrafast and memory optimized gene set enrichment scoring algorithm not unlike fGSEA. ultragsea demonstrates highly correlated gene set scores and p-values and typically is 100 times faster than fGSEA.

# Example

## Preparing gensets

For this vignette, we retrieve the Hallmark gene sets using the `msigdbr` package. Of course you can use your own gene sets or download from somewhere else.

```{r}
library(ultragsea)
library(msigdbr)
gs <- msigdbr::msigdbr(collection = "H")
gmt <- tapply(gs$gene_symbol,gs$gs_name,list)
length(gmt)
```

## Performing the statistical test

Just for this example we use some random values for `fc` but in general this is a fold change from some statistical comparison of your experiment. 

```{r}
G <- gmt2mat(gmt)
fc <- rnorm(nrow(G))
names(fc) <- rownames(G)
res <- ultragsea(G, fc, format='as.gsea', method='cor')
res <- ultragsea(G, fc, format='as.gsea', method='ztest')
head(res)
```

For the fastest speed, especially when computing enrichment for multiple foldchanges, it is better to call the low-level functions for geneset correlation or geneset z-test directly. They do not return the nicely formatted output tables as `ultragsea()` but they are much faster as they are optimized for matrix inputs `F`.

```{r}
F <- cbind(fc,fc,fc,fc,fc)
res1 <- gset.cor(G, F, compute.p=TRUE)
res2 <- gset.ztest(G, F)
```


## Plotting

You can proceed by plotting the gene set using e.g. the plotting functions of the fgsea package.

```{r}
#library(fgsea)
#k <- res$pathway[which.min(res$pval)]
#fgsea::plotEnrichment(gmt[[k]],fc,ticksSize = 0.2)
```



# Session info

```{r}
sessionInfo()
```
